---
title: "M05: Reflection on Ch09 (Evaluation)"
author: "Grace Sen"
date: today
format:

  html:
    toc: true
    toc-depth: 4
    toc-expand: 1
    toc-location: right-body
    toc-title: "Contents"
    number-sections: true
    code-fold: show
    code-tools: true
    theme: cosmo
    highlight-style: github
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
  freeze: true
---

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(tidymodels)
library(tidyverse)
library(gt)
library(patchwork)

tidymodels_prefer()

set.seed(2026)

# Synthetic marketing dataset
set.seed(123)
n <- 3000

mail_data <- tibble(
  customer_id = paste0("C", str_pad(1:n, 5, pad = "0")),
  age = round(rnorm(n, mean = 45, sd = 12)),
  income = round(rlnorm(n, meanlog = 10.8, sdlog = 0.6)),
  recency_days = round(rexp(n, rate = 1 / 60)),
  freq_12mo = rpois(n, lambda = 3),
  avg_order_amt = round(rlnorm(n, meanlog = 4.2, sdlog = 0.5), 2),
  channel = sample(
    c("email", "direct_mail", "digital"),
    n,
    replace = TRUE,
    prob = c(0.5, 0.3, 0.2)
  ),
  region = sample(
    c("West", "South", "Midwest", "Northeast"),
    n,
    replace = TRUE
  ),
  loyalty_tier = factor(
    sample(
      c("Bronze", "Silver", "Gold", "Platinum"),
      n,
      replace = TRUE,
      prob = c(0.4, 0.3, 0.2, 0.1)
    ),
    levels = c("Bronze", "Silver", "Gold", "Platinum")
  ),
  responded = rbinom(
    n,
    1,
    prob = plogis(
      -3 +
        0.02 * (age - 45) +
        0.3 * log(income / 50000) +
        0.1 * freq_12mo -
        0.005 * recency_days
    )
  )
) |>
  mutate(
    income = if_else(runif(n) < 0.08, NA_real_, income),
    avg_order_amt = if_else(runif(n) < 0.05, NA_real_, avg_order_amt),
    responded = factor(responded,
      levels = c(1, 0),
      labels = c("yes", "no")
    )
  )

# Train/test split
set.seed(617)

mail_split <- initial_split(
  mail_data,
  prop = 0.80,
  strata = responded
)

mail_train <- training(mail_split)
mail_test <- testing(mail_split)

# Recipe
mail_rec <- recipe(
  responded ~ age +
    income +
    recency_days +
    freq_12mo +
    avg_order_amt +
    channel +
    region +
    loyalty_tier,
  data = mail_train
) |>
  step_impute_median(all_numeric_predictors()) |>
  step_log(
    income,
    avg_order_amt,
    recency_days,
    base = 10,
    offset = 1
  ) |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())

# Model
lr_spec <- logistic_reg(
  penalty = 0.01,
  mixture = 1
) |>
  set_engine("glmnet") |>
  set_mode("classification")

mail_wf <- workflow() |>
  add_recipe(mail_rec) |>
  add_model(lr_spec)

mail_fit <- fit(mail_wf, data = mail_train)

test_preds <- augment(mail_fit, new_data = mail_test)
```

# **Exercise 9.1 — Confusion Matrix and Threshold Tradeoff**

## **Part A — Reading the Confusion Matrix**

> *Run `conf_mat(test_preds, truth = responded, estimate = .pred_class)` and use the output to answer:*

### *How many customers actually responded (“yes”) and were correctly predicted to respond? What is this cell called?*

::: callout-note
# Answer

**0 customers.** This cell is called the **true positive (TP)**.
:::

### *How many customers actually responded but were predicted “no”? What is this cell called, and what does it cost a marketer?*

::: callout-note
# Answer 

**31 customers.** This is a **false negative (FN)**. It means the business missed customers who would have responded, resulting in lost sales opportunities.
:::

### *How many customers did not respond but were predicted “yes”? What is this cell called, and what does it cost a marketer?*

::: callout-note
# Answer

**0 customers.** This is a **false positive (FP)**. It would waste money by sending marketing to customers who would not respond.
:::

### *In plain English, which type of error is more costly for a direct-mail campaign with a \$5 mailing cost and \$50 average order value — a false positive or a false negative? Explain your reasoning.*

::: callout-note
## Answer

A **false negative** is more costly because missing a customer who would have purchased loses much more potential revenue than the \$5 mailing cost.
:::

```{r}
conf_mat(test_preds,
         truth = responded,
         estimate = .pred_class)
```

## **Part B — Threshold Tradeoff**

> *Using `test_preds`, compute `precision` and `sens` at the default 0.5 threshold. Now manually create a new `.pred_class` column using a threshold of 0.3 instead (classify as “yes” if `.pred_yes > 0.3`). Recompute both metrics. Which one increased, and why does lowering the threshold have that effect?*

::: callout-note
## Answer

Lowering the threshold to **0.3** did **not** change the results. The model still predicted every customer as **"no"**, so **precision remained NA** and **sensitivity remained 0** because no predicted probabilities exceeded 0.3.
:::

```{r}
# Default threshold (0.5)
metric_set(precision, sens)(
  test_preds,
  truth = responded,
  estimate = .pred_class
)

# New threshold (0.3)
test_preds_03 <- test_preds |>
  mutate(
    .pred_class = factor(
      if_else(.pred_yes > 0.3, "yes", "no"),
      levels = levels(responded)
    )
  )

metric_set(precision, sens)(
  test_preds_03,
  truth = responded,
  estimate = .pred_class
)
```

# **Exercise 9.2 — ROC/AUC Interpretation and Custom `metric_set()`**

## **Part A — Interpreting AUC**

> *The notebook computed `roc_auc` for the direct-mail response model.*

### *Look up the AUC value from the `roc-curve` chunk. Is it above or below 0.65 (the “usable for targeting” threshold from the callout)?*

::: callout-note
# Answer

The AUC is **0.473**, which is **below 0.65**. This means the model is not good enough for targeting customers.
:::

### *In plain English, what does an AUC of that value mean? Complete this sentence: “If we randomly pick one customer who responded and one who did not, the model assigns a higher predicted probability to the responder \_\_\_\_% of the time.”*

::: callout-note
# Answer

An AUC of **0.473** means the model assigns a higher predicted probability to a customer who responded only **47.3%** of the time. This is worse than random guessing.
:::

### *The ROC curve plot shows the sensitivity/specificity tradeoff across all thresholds. At what point on the curve would you operate if your campaign budget only allows you to mail the top 10% of prospects?*

::: callout-note
## Answer

If the budget only allows mailing the top 10% of customers, I would choose a point on the ROC curve with a **low false positive rate** while keeping the **highest possible sensitivity**.
:::

## **Part B — Custom `metric_set()`**

> *Create a `metric_set()` containing `accuracy`, `roc_auc`, `sens`, `spec`, and `f_meas`. Apply it to `test_preds`. If you had to report just **two** of these five numbers to a VP of Marketing who is unfamiliar with machine learning metrics, which would you choose and why?*

::: callout-note
## Answer

I would report **ROC AUC** and **sensitivity**. ROC AUC shows how well the model separates responders from non-responders, while sensitivity tells us how many actual responders the model identifies. These metrics are more useful than accuracy for an imbalanced dataset.
:::

```{r}
my_metrics <- metric_set(
  accuracy,
  roc_auc,
  sens,
  spec,
  f_meas
)

test_preds |>
  my_metrics(
    truth = responded,
    estimate = .pred_class,
    .pred_yes
  )
```

# **Exercise 9.3 — Regression Metric Sensitivity to Outliers**

> *Add one synthetic “whale” customer to `reg_preds` with `order_value_next = 5000` and `.pred = 200` (a badly missed prediction). Recompute `rmse` and `mae`. Which metric moved more in percentage terms? What does this tell you about choosing RMSE vs. MAE for marketing data with occasional huge orders?*

::: callout-note
## Answer

**RMSE** increased much more than **MAE** after adding the whale customer. This shows that RMSE is more sensitive to large prediction errors, while MAE is more robust to outliers.
:::

```{r}
#| echo: false
#| warning: false
#| message: false

mail_data <- mail_data |>
  mutate(
    order_value_next = round(
      pmax(
        0,
        15 +
          0.0006 * coalesce(income, median(income, na.rm = TRUE)) +
          4 * freq_12mo -
          0.05 * recency_days +
          rnorm(n(), 0, 20)
      ),
      2
    )
  )

set.seed(617)

mail_split <- initial_split(
  mail_data,
  prop = 0.80,
  strata = responded
)

mail_train <- training(mail_split)
mail_test <- testing(mail_split)

reg_rec <- recipe(
  order_value_next ~ income + freq_12mo + recency_days,
  data = mail_train
) |>
  step_impute_median(all_numeric_predictors()) |>
  step_normalize(all_numeric_predictors())

reg_spec <- linear_reg() |>
  set_engine("lm")

reg_wf <- workflow() |>
  add_recipe(reg_rec) |>
  add_model(reg_spec)

reg_fit <- fit(reg_wf, data = mail_train)

reg_preds <- augment(reg_fit, new_data = mail_test)
```

```{r}
# Add one extreme customer
reg_preds_whale <- bind_rows(
  reg_preds,
  tibble(
    customer_id = "Whale",
    order_value_next = 5000,
    .pred = 200,
    .resid = 5000 - 200
  )
)

# Recalculate metrics
metric_set(rmse, mae)(
  reg_preds_whale,
  truth = order_value_next,
  estimate = .pred
)
```

# **Exercise 9.4 — Multiclass Averaging**

> *Re-run the multiclass `loyalty_tier` example using `estimator = "micro"` instead of the default macro averaging for `sens`. Also run it with `estimator = "macro_weighted"`. Compare all three numbers.*

### *Why might micro-averaging look more favorable when one class (Bronze) is much larger than the others?*

::: callout-note
# Answer

Micro-averaging looks better because it is influenced more by the larger Bronze class.
:::

### *Which estimator would you choose if you wanted to ensure that **Platinum** customers — the rarest and most valuable tier — are not underrepresented in the metric? Explain your reasoning.*

::: callout-note
# Answer

I would choose **macro averaging** because it gives equal importance to every loyalty tier, including the rare Platinum customers.
:::

### *In a marketing context, which averaging method is most likely to give a misleadingly optimistic picture of model performance? Why?*

::: callout-note
## Answer

**Micro averaging** is the most misleading because it is dominated by the largest class and can make the model look better than it is.
:::

```{r}
#| echo: false
#| warning: false
#| message: false

multi_rec <- recipe(
  loyalty_tier ~ age + income,
  data = mail_train
) |>
  step_impute_median(all_numeric_predictors()) |>
  step_normalize(all_numeric_predictors())

multi_spec <- multinom_reg() |>
  set_engine("nnet")

multi_wf <- workflow() |>
  add_recipe(multi_rec) |>
  add_model(multi_spec)

multi_fit <- fit(multi_wf, data = mail_train)

multi_preds <- augment(multi_fit, new_data = mail_test)
```

```{r}
# Macro averaging
multi_preds |>
  metric_set(accuracy, sens)(
    truth = loyalty_tier,
    estimate = .pred_class,
    estimator = "macro"
  )

# Micro averaging
multi_preds |>
  metric_set(accuracy, sens)(
    truth = loyalty_tier,
    estimate = .pred_class,
    estimator = "micro"
  )

# Macro-weighted averaging
multi_preds |>
  metric_set(accuracy, sens)(
    truth = loyalty_tier,
    estimate = .pred_class,
    estimator = "macro_weighted"
  )
```

# Appendix

GitHub Repo:\
GitHub Pages:
